---
title: "Smart Typer"
author: "KOSW"
date: "2024-02-25"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```




## Problem to solve

These days, typing with PC or smartphone is core means of communication.
But while we're typing, it is inefficient to type all alphabets one-by-one.

So, it is required to improve the efficiency.


## If Problem is solved?

- Saving time by typing same contents faster.
- Helping elderly people who are unfamiliar with typing by recommending the most likely words.

## How to solve?

- Statistical method

```{r, echo=FALSE, message=FALSE}
if(!("png" %in% installed.packages())){
  install.packages('png')
}

library(png)
```


```{r, fig.height=3, fig.pos='t'}
#image
x <- "https://raw.githubusercontent.com/rejeton1/Data_Science_Capstone_Project/main/How%20to%20solve%20img.png"
download.file(x,'howtosolve.png', mode = 'wb')
pp1 <- readPNG("howtosolve.png")
plot.new() 
rasterImage(pp1,0,0,1,1)
```

- Based on previous two words, recommend 3 words that has the biggest probability.
- If previous two words don't exist, based on previous one word.
- If that's not the case, it recognize it is a starting poing of sentence.

* How can it recommend OOV(Out-Of-Vocabulary) in train set?
It can recommend them by giving them a small probability through smoothing method.

## How it works?

```{r, fig.height=7, fig.width=7, fig.align='center'}
#image
y <- "https://raw.githubusercontent.com/rejeton1/Data_Science_Capstone_Project/main/howwork.png"
download.file(y,'howwork.png', mode = 'wb')
pp2 <- readPNG("howwork.png")
plot.new() 
rasterImage(pp2,0,0,1,1)
```


## Performance

```{r}
#image
z <- "https://raw.githubusercontent.com/rejeton1/Data_Science_Capstone_Project/main/performance.png"
download.file(z,'perform.png', mode = 'wb')
pp3 <- readPNG("perform.png")
plot.new() 
rasterImage(pp3,0,0,1,1)
```
When type above document with/without this app,
How long did it take?

1. typing one-by-one :     2. typing with app : 


Prediction Accuracy : 0.138
(about test set size 1852)
