---
title: "Smart Typer"
author: "KOSW"
date: "2024-02-25"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Problem to solve

These days, typing with PC or smartphone is core means of communication.
But while we're typing, it is inefficient to type all alphabets one-by-one.

So, it is required to improve the efficiency.

### If Problem is solved?

- Saving time by typing same contents faster.
- Helping elderly people who are unfamiliar with typing by recommending the most likely words.

## How to solve?

- Statistical method

```{r, echo=FALSE}
# Define variable containing url
url1 <- "https://raw.githubusercontent.com/rejeton1/Data_Science_Capstone_Project/main/How%20to%20solve%20img.png"
```

<center>![](`r url1`){width=40%, height=40%}</center>

1. Based on previous two words, recommend 3 words that has the biggest probability.
2. If previous two words don't exist, based on previous one word.
3. If that's not the case, it recognize it is a starting poing of sentence.

## How it works?
```{r, echo=FALSE}
# Define variable containing url
url2 <- "https://raw.githubusercontent.com/rejeton1/Data_Science_Capstone_Project/main/howwork.png"
```

<center>![](`r url2`){width=65%}</center>

## Faster and Accurate

```{r, echo=FALSE}
# Define variable containing url
url3 <- "https://raw.githubusercontent.com/rejeton1/Data_Science_Capstone_Project/main/performance.png"
```

<center>![](`r url3`)</center>

### Performance 1. Speed

When type above document with/without this app,
How long did it take?

1. typing one-by-one :     2. typing with app : 

### Performance 2. Accuracy

Prediction Accuracy : 0.138
(about test set size 1852)

